---
title: "relational_data"
author: "Brendan Doyle"
date: "October 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library ("tidyverse")
library ("nycflights13")
library ("Lahman")
library ("babynames")
library ("nasaweather")
library ("fueleconomy")
library ("ggplot2")
library ("maps")
library ("viridis")
```

airlines
airports
planes
flights
weather

### Exercises 174
1. Use the flight table because it has the origin and dest variables. Use the airports table as well for longitude and latitude of the flight, this table as has dest and origin. Combine flights with airports to get the "origin" airport and to get to "dest" airport location.

2. Their relationship is with the "origin" variable in weather and with "faa" in airports

3. The year, month, day, hour, and origin in weather needs to paired with the same variable names in flight, plus dest. 

4. Create a new table for special dates. The main variables would be year, month, and day and matched with the ones of flights table.

###Exercises 177

```{r ex 177 Q1}
flights %>% 
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_surrogate = row_number()) %>%
  glimpse()
```
?glimpse()
1. above

Lahman::Batting
babynames::babynames
nasaweather::atmos
fueleconomy::vehicles
ggplot2::diamonds


```{r ex 177 Q2}
Lahman::Batting %>%
  group_by(playerID, yearID, stint)

babynames::babynames %>%
  group_by(year, sex, name)

nasaweather::atmos %>%
  group_by(lat, long, year, month)

fueleconomy::vehicles %>%
  group_by(id)

ggplot2::diamonds %>%
  distinct() %>%
  nrow()
```

2. above

3. primary key of "playerID, yearID, stint", for Batting and the foreign key of "playerID" can join with the same named primary key in Master table. Salaries table shares a primary key of PlayerID with Batting and Master, and a primary key of yearID and teamID with the primary keys of Managers. Then managers table shares the same primary key with AwardsManagers table with yearID and playerID. 

```{r mutating joins}
flights2 <- flights %>%
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2 %>%
  select(-origin, -dest) %>%
  left_join(airlines, by = "carrier")
```

###Exercises 186

```{r ex 186 Q1}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()

avg_dest_delays <-
  flights %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa"))

avg_dest_delays %>%
  ggplot(aes(lon, lat, colour = delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()

```
1.  above



2. 

```{r question 2 pg187}
flights %>%
  left_join(airports, by = c(dest = "faa")) %>%
  left_join(airports, by = c(origin = "faa")) %>%
  head()

```

3. There doesn't appear to be a relationship between age of a plane and its delays as should by the graph below. 

```{r Q3 187}
plane_ages <- 
  planes %>%
  mutate(age = 2013 - year) %>%
  select(tailnum, age)

flights %>%
  inner_join(plane_ages, by = "tailnum") %>%
  group_by(age) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean(dep_delay)) %>%
  ggplot(aes(x = delay, y = age)) +
  geom_point() +
  geom_line()


```

4. The weather conditions, specifically rain precipitation relates to a delay. However you would expect most of a delay with more precipitation but it is strongest with around 0.2 because it is probably more likely to rain that amount more often at 1.2 as shown by the amount of plotted points.

```{r Q4 187}
flight_weather <-
  flights %>%
  inner_join(weather, by = c("origin" = "origin",
                            "year" = "year",
                            "month" = "month",
                            "day" = "day",
                            "hour" = "hour"))

flight_weather %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = precip, y = delay)) +
    geom_line() + geom_point()

```

5. Google showed that the southeast has a lot of visible weather conditions, specifically storms. The yellow indicates where the biggest delay was which based on the latitude and longitude is 35 (ish) degrees and -85 degrees, respectively. It looks like on the map that it is in Tennessee.
```{r Q5}
library(viridis)


flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()  + 
  scale_color_viridis()

```

### Filtering Joins











